Parsing Tweets into Universal Dependencies

### 2018-04-04 paper.tex
feedback on recent edits

@@

\subsection{Comparison to PoSTWITA-UD}\label{sec:postwita}
\citet{sanguinetti-17} presented the \textit{PosTWITA-UD} which is
the first Twitter treebank in Universal Dependencies.
In their convention, tokenization tends to preserve the original
tweet content but two special cases, \textit{articulated prepositions}
and \textit{clitic clusters}, are tokenized.
Beyond tokenization, an extra layer of lemmas is annotated
to recover ``what is understandable by a human''.
It's arguable that normalization (the lemmatization in \citet{sanguinetti-17})
on social media data is essential according to \citet{eisenstein:2013:NAACL-HLT}.
In this paper, we don't include normalization and leave it an open question for further study.

In the syntax annotation, tweet specific
constructions are considered by both \citet{sanguinetti-17} and our paper.
We agree on the syntax head in these constructions but differ on
the choice of relations for depicting these constructions.
We use \textit{discourse} for topical hashtag and \citet{sanguinetti-17} used \textit{parataxis}. 
Similar difference happens in referential URLs where we use \textit{list} by following UD\_English and they use \textit{dep}.
The choice of relation on the non-syntactic tokens is questionable
since UD doesn't have specified guideline for this. 
Our choice of \textit{discourse} is inspired by the observation that emoticon is annotated as \textit{discourse}. Non-syntactic tokens can be easily recovered
from our annotation and changing it to other relation in the future development is not difficult.

@@

\nss{discuss similarities and differences of conventions to \cite{sanguinetti-17}. 
We expect that the official UD guidelines will eventually adopt standards for these issues so the treebanks can be harmonized.}
\yjcomment{i've finished this part.}

@@

\subsubsection{Annotation Process}
\begin{table}
	\centering
	\begin{tabular}{rrrrr}
		\hline
		 & \multicolumn{2}{c}{\textsc{Tweebank v1}} & \multicolumn{2}{c}{\textsc{Tweebank v2}} \\
		split & tweets & tokens & tweets & tokens \\
		\hline
		train & 639 & 9,310 & 1,639 & 24,753 \\
		dev & - & - & 709 & 11,742 \\
		test & 201 & 2,839 & 1,201 & 19,112 \\
		\hline
		total & 840 & 12,149 & 3,549 & 55,617 \\
%		\textsc{v1} & tweets & 639 & 0 & 201 & 840 \\
%		& tokens  & 9,310 & 0 & 2,839 & 12149 \\
%		\hdashline
%		\textsc{v2} & tweets & 1,639 & 709 & 1,201 & 3,549 \\
%		 & tokens &	24,753	&	11,742	&	19,112 & 55,617\\
		\hline
	\end{tabular}

%	\begin{tabular}{rrrrrr}
%		\hline
%		\multicolumn{2}{l}{version} & train & dev & test & total \\
%		\hline
%		\textsc{v1} & tweets & 639 & 0 & 201 & 840 \\
%		& tokens  & 9,310 & 0 & 2,839 & 12149 \\
%		\hdashline
%		\textsc{v2} & tweets & 1,639 & 709 & 1,201 & 3,549 \\
%		 & tokens &	24,753	&	11,742	&	19,112 & 55,617\\
%		\hline
%	\end{tabular}
\caption{Statistics of {\sc Tweebank v2} and the comparison with
{\sc Tweebank v1} in quantity.\nss{not obvious how to interpret this: train tokens looks like 1,609+1,000 = 2,609, but maybe it's 609+1,000=1,609? 
I think it would be easier just to show v1 and v2 as separate rows. Also, a column for totals across the splits.}
\yjcomment{i've fix this.}
\label{tbl:data-stat}
}
\end{table}
Our annotation process was conducted in two stages.
In the first stage, 18 researchers worked on the {\sc Tweebank v1}
proportion and the additional 210 tweets and created the initial annotations in one day.
Before annotating, they were given a tutorial overview of the general UD
annotation conventions and our guidelines specifically for annotating tweets.
Both the guidelines and annotations
were further refined by the authors of this paper to increase
the coverage of our guidelines and solve inconsistencies between
different annotators during this exercise. In the second stage, a tokenizer, a POS tagger, and a
parser were trained on the annotated data from the first stage (1,050 tweets in total),
and used to automatically analyze the sampled 2,500 tweets.  Authors 
of this paper manually corrected the parsed data and finally achieved 3,550 labeled tweets\nss{What tooks did annotators use?}\yjcomment{fixed}.\footnote{We use arborator \cite{gerdes:2013:W13-37} in our annotation.}
Newly created annotations are split into train, development, and test set and appended
to the original split of {\sc Twebank v1}. Statistics of our annotations and data split are shown
in Table \ref{tbl:data-stat}.

@@

\begin{table}[t]
	\centering
	\begin{tabular}{rc}
		\hline
		\it System & $F_1$ \\
		\hline
		  Stanford CoreNLP & 96.6 \\
		 Twokenizer & 94.4 \\		 
% 		UDPipe v1.2 & 94.7 \\
		\hdashline
		UDPipe v1.2 & 97.3 \\
	 	our bi-LSTM tokenizer & 98.3 \\
		\hline
	\end{tabular}
	\caption{Tokenizer comparison on the \textsc{Tweebank v2} test set.}\label{tbl:tok-result}
\end{table}
\citet{wang-EtAl:2017:Long6} show that the parser\nss{tokenizer?}\yjcomment{it's a parsing work.} trained 
on the combination of the UD\_English and Singapore English dialect (Singlish) treebanks
outperforms the one trained only on the dialect treebank.  Preliminary results
showed the same in our case, so 
we follow their work and trained our tokenizer on the training portion of
\textsc{Tweebank v2} combined with the UD\_English training set
and tested on the \textsc{Tweebank v2} test set. % \yicomment{should mention that combination of corpora improves the results, otherwise we should answer why we didn't try training ONLY on Tweebank v2}\yjcomment{add discussion}
We report $F_1$ scores, combining precision and recall for token identification. Table \ref{tbl:tok-result} shows the
tokenization results, compared to  other available tokenizers. 
 Stanford CoreNLP \cite{manning-EtAl:2014:P14-5} and Twokenizer
\cite{ICWSM101540} are rule-based systems and were not adapted
to the UD tokenization scheme.
The UDPipe v1.2
\cite{straka-strakova:2017:K17-3} model was re-trained on the same
data as our system. Compared with UDPipe, we use LSTM
instead of GRU in our model and we also use a larger size of hidden units (64 vs.~20),
which has stronger representation power. 
Our bi-LSTM tokenizer achieves the best accuracy among all these
tokenizers.  These results indicate the value of statistical modeling
in tokenization for informal texts.\nss{this paragraph needs copyediting}

@@

\section*{Acknowledge}
We thank Elizabeth Clark, Lucy Lin, Nelson Liu, Kelvin Luu, Phoebe Mulcaire, 
Hao Peng, Maarten Sap, Chenhao Tan, Sam Thomson,
and the other colleagues in Georgetown University for the annotation efforts in the first round.
We also thank the anonymous  reviewers for their helpful comments and suggestions.
\yjcomment{fonding agency here and anyone else i miss.}
This work was supported by the National Key Basic Research
Program of China via grant 2014CB340503 and the
National Natural Science Foundation of China (NSFC) via
grant 61632011.

>>>>

\newcommand{\citeposs}[2][]{\citeauthor{#2}'s (\citeyear[#1]{#2})}
\newcommand{\Citeposs}[2][]{\Citeauthor{#2}'s (\citeyear[#1]{#2})}

@@

\subsection{Comparison to PoSTWITA-UD}\label{sec:postwita}
The first Twitter treebank annotated with Universal Dependencies 
was the \mbox{\textit{PosTWITA-UD}} corpus for Italian \citep{sanguinetti-17}, 
which consists of \nss{??} tweets (\nss{??} sentences, \nss{??} tokens).
In their convention, tokenization tends to preserve the original
tweet content but two special cases, \textit{articulated prepositions}
and \textit{clitic clusters}, are tokenized.\nss{explain what these two phenomena are. are these the only words that are split?}
Beyond tokenization, an extra layer of lemmas is annotated
to recover ``what is understandable by a human''.
It's arguable that normalization (the lemmatization in \citet{sanguinetti-17})
on social media data is essential according to \citet{eisenstein:2013:NAACL-HLT}.\nss{``It's arguable that normalization...is essential'' suggests that we think it IS essential. Did you mean ``whether'' instead of ``that''?}
In this paper, we don't include normalization and leave it an open question for further study.\nss{Maybe just say: 
Their lemmas include spelling normalization, whereas our lemmas only normalize casing and inflectional morphology (if that is correct). 
You could also mention that the current UD guidelines on lemmas (which I wrote) give a lot of flexibility to treebank 
developers: \url{http://universaldependencies.org/u/overview/morphology.html\#lemmas}}

%In the syntax annotation, tweet specific
%constructions are considered by both \citet{sanguinetti-17} and our paper.
With respect to tweet-specific constructions,
our and \citeposs{sanguinetti-17} interpretation of headedness is the same, 
but we differ in the relation label.
For topical hashtags, we use \textit{discourse} while \citet{sanguinetti-17} used \textit{parataxis}. 
In referential URLs, we use \textit{list} (following the precedent of UD\_English-EWT) while \citet{sanguinetti-17} used \textit{dep}.
The choice of relation on the non-syntactic tokens is questionable
since UD doesn't specify a convention for this. 
Our choice of \textit{discourse} is inspired by the observation that emoticon is annotated as \textit{discourse}.\nss{what did PosTWITA-UD use for non-syntactic tokens?} 
% Non-syntactic tokens can be easily recovered
% from our annotation and changing it to other relation in the future development is not difficult.
We expect that the official UD guidelines will eventually adopt standards for these constructions so the treebanks can be harmonized.

@@

@@

\subsubsection{Annotation Process}
\begin{table}
	\centering
	\begin{tabular}{rrrrr}
		\hline
		 & \multicolumn{2}{c}{\textsc{Tweebank v1}} & \multicolumn{2}{c}{\textsc{Tweebank v2}} \\
		split & tweets & tokens & tweets & tokens \\
		\hline
		train & 639 & 9,310 & 1,639 & 24,753 \\
		dev & --\hphantom{0} & --\hphantom{00} & 709 & 11,742 \\
		test & 201 & 2,839 & 1,201 & 19,112 \\
		\hline
		total & 840 & 12,149 & 3,549 & 55,617 \\
%		\textsc{v1} & tweets & 639 & 0 & 201 & 840 \\
%		& tokens  & 9,310 & 0 & 2,839 & 12149 \\
%		\hdashline
%		\textsc{v2} & tweets & 1,639 & 709 & 1,201 & 3,549 \\
%		 & tokens &	24,753	&	11,742	&	19,112 & 55,617\\
		\hline
	\end{tabular}

%	\begin{tabular}{rrrrrr}
%		\hline
%		\multicolumn{2}{l}{version} & train & dev & test & total \\
%		\hline
%		\textsc{v1} & tweets & 639 & 0 & 201 & 840 \\
%		& tokens  & 9,310 & 0 & 2,839 & 12149 \\
%		\hdashline
%		\textsc{v2} & tweets & 1,639 & 709 & 1,201 & 3,549 \\
%		 & tokens &	24,753	&	11,742	&	19,112 & 55,617\\
%		\hline
%	\end{tabular}
\caption{Statistics of {\sc Tweebank v2} and the comparison with
{\sc Tweebank v1} in quantity.
%\nss{not obvious how to interpret this: train tokens looks like 1,609+1,000 = 2,609, but maybe it's 609+1,000=1,609? 
%I think it would be easier just to show v1 and v2 as separate rows. Also, a column for totals across the splits.}
%\yjcomment{i've fix this.}
\label{tbl:data-stat}
}
\end{table}
Our annotation process was conducted in two stages.
In the first stage, 18 researchers worked on the {\sc Tweebank v1}
proportion and the additional 210 tweets and created the initial annotations in one day.
Before annotating, they were given a tutorial overview of the general UD
annotation conventions and our guidelines specifically for annotating tweets.
Both the guidelines and annotations
were further refined by the authors of this paper to increase
the coverage of our guidelines and solve inconsistencies between
different annotators during this exercise. In the second stage, a tokenizer, a POS tagger, and a
parser were trained on the annotated data from the first stage (1,050 tweets in total),
and used to automatically analyze the sampled 2,500 tweets.  Authors 
of this paper manually corrected the parsed data and finally achieved 3,550 labeled tweets.\footnote{Manual annotation was done with Arborator \cite{gerdes:2013:W13-37}, a web platform for drawing dependency trees.}
Newly created annotations are split into train, development, and test set and appended
to the original split of {\sc Twebank v1}. Statistics of our annotations and data split are shown
in Table \ref{tbl:data-stat}.

@@

\begin{table}[t]
	\centering
	\begin{tabular}{rc}
		\hline
		\it System & $F_1$ \\
		\hline
		  Stanford CoreNLP & 96.6 \\
		 Twokenizer & 94.4 \\		 
% 		UDPipe v1.2 & 94.7 \\
		\hdashline
		UDPipe v1.2 & 97.3 \\
	 	our bi-LSTM tokenizer & 98.3 \\
		\hline
	\end{tabular}
	\caption{Tokenizer comparison on the \textsc{Tweebank v2} test set.}\label{tbl:tok-result}
\end{table}
\citet{wang-EtAl:2017:Long6} show that the parser\nss{tokenizer?}\yjcomment{it's a parsing work.}\nss{I guess I don't follow how parsing results 
on combined training data affect our decision about how to tokenize.} trained 
on the combination of the UD\_English and Singapore English dialect (Singlish) treebanks
outperforms the one trained only on the dialect treebank.  Preliminary results
showed the same in our case, so 
we follow their work and trained our tokenizer on the training portion of
\textsc{Tweebank v2} combined with the UD\_English training set
and tested on the \textsc{Tweebank v2} test set. % \yicomment{should mention that combination of corpora improves the results, otherwise we should answer why we didn't try training ONLY on Tweebank v2}\yjcomment{add discussion}
We report $F_1$ scores, combining precision and recall for token identification. Table \ref{tbl:tok-result} shows the
tokenization results, compared to  other available tokenizers. 
 Stanford CoreNLP \cite{manning-EtAl:2014:P14-5} and Twokenizer
\cite{ICWSM101540} are rule-based systems and were not adapted
to the UD tokenization scheme.
The UDPipe v1.2
\cite{straka-strakova:2017:K17-3} model was re-trained on the same
data as our system. Compared with UDPipe, we use LSTM
instead of GRU in our model and we also use a larger size of hidden units (64 vs.~20),
which has stronger representation power. 
Our bi-LSTM tokenizer achieves the best accuracy among all these
tokenizers.  These results indicate the value of statistical modeling
in tokenization for informal texts.\nss{this paragraph needs copyediting}

@@

\section*{Acknowledge}
We thank Elizabeth Clark, Lucy Lin, Nelson Liu, Kelvin Luu, Phoebe Mulcaire, 
Hao Peng, Maarten Sap, Chenhao Tan, and Sam Thomson at the University of Washington, 
and Austin Blodgett, Lucia Donatelli, Joe Garman, Emma Manning, Angela Yang, and Yushi Zhang 
at Georgetown University for their annotation efforts in the first round.
We also thank the anonymous reviewers for their helpful comments and suggestions.
\yjcomment{funding agency here and anyone else i miss.}
This work was supported by the National Key Basic Research
Program of China via grant 2014CB340503 and the
National Natural Science Foundation of China (NSFC) via
grant 61632011.
